Future Parameter constraints from weak lensing CMB and Galaxy Lensing Power- and Bi-spectra

### 2025-07-11 paper/paper_draft.tex
Updates from Overleaf

\usepackage[utf8]{inputenc}   % Allows UTF-8 character encoding
\usepackage{amsmath}          % For advanced math typesetting
\usepackage{amssymb}          % Additional math symbols
\usepackage{graphicx}         % For including images
\usepackage{hyperref}         % For hyperlinks in the document
\usepackage{geometry}         % For setting up page geometry
\geometry{margin=1in}         % 1-inch margins all around
\usepackage{siunitx}          % For consistent typesetting of units
\usepackage{float}            % To control float positioning (e.g., figures/tables)
\usepackage{enumitem}         % Better control over list formatting
\usepackage{tocloft}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{shapes.geometric, arrows, backgrounds}
\usepackage{empheq}
\usepackage{booktabs}
\usepackage{multirow} % needed for merging cells vertically
\usepackage{booktabs}

@@

\emailAdd{jonasfrugte@gmail.com}
\emailAdd{p.d.meerburg@rug.nl}
\abstract{
Upcoming Stage 4 surveys, such as the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision. While the power spectrum is the standard statistic used to analyze weak lensing data, non-Gaussianity induced by the non-linear growth of structure imparts significant cosmological information into higher-order statistics. In this paper, we forecast the ability of future surveys to constrain cosmological parameters using the weak lensing power spectrum and bispectrum from both CMB and galaxy surveys, as well as their cross-correlations. We consider an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for Stage 4 survey specifications. Our analysis shows that, in the absence of systematics, both the CMB and galaxy lensing bispectra will be detectable at high signal-to-noise. We consider parameter constraints with a ``weak'' prior and with a stronger prior from the primary CMB anisotropies. In the weak prior case the lensing bispectrum substantially improves parameter constraints over using the lensing power spectrum alone, primarily by breaking parameter degeneracies. For the stronger prior we see less significant gains, particularly for the CMB lensing bispectrum. We further find that for stage 4 surveys that are able to measure lensing on small scales where non-linear structure formation becomes important, the constraining power of the lensing bispectrum can become comparable to that of the lensing power spectrum. Furthermore, we demonstrate that, both for our weak and strong priors, the synergy between CMB and galaxy lensing is substantial. Combining the two probes, particularly with both their lensing power- and bi-spectra, leads to significantly tighter constraints than achievable with either probe individually, with the greatest improvement seen for the sum of neutrino masses.
}

@@

Weak lensing has been used as a probe to constrain cosmological parameters for over a decade \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001,Planck2018Lensing}. A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aims to measure weak lensing of the cosmic microwave background (CMB) and galaxies respectively. These upcoming surveys are expected to achieve significantly higher accuracy than their predecessors, leading to tighter constraints on cosmological parameters and advancing our overall understanding of cosmology. Typically, the standard summary statistic of interest is the power spectrum (or equivalently, two-point function) of the lensing potential (as well as cross-correlations with galaxy clustering). This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$. In the weak lensing regime and under the Born approximation they are all equivalent, in the sense that they can be directly converted into one another (appendix \ref{sec:weaklensstats}). In CMB surveys, the temperature and polarization anisotropies are used to reconstruct the lensing potential \cite{Hu_2002}, while with galaxies one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipticity distribution of the observed galaxies \cite{HoekstraJain2008}.

Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}. In addition to looking at the power spectrum, a natural next step is thus to measure weak lensing bispectra. While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending \cite{Kalaja_2023,Namikawa_2016}. Upcoming surveys, such as the ones mentioned earlier may be able to detect this signal. Therefore it is timely to consider the potential of higher order spectra to constrain cosmological parameters.

@@

\subsection{Parameter Constraints}
Next, we consider cosmological parameter constraints. We include both the power spectrum and the bispectrum of the weak lensing. We look at two extension of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as the derived parameters $\sigma_8$, and $\Omega_m$. Constraints on the full set of $\Lambda$CDM parameters can be found in appendix \ref{sec:lambdacdmconstraints}. Results shown include marginalization over all parameters. Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses from stage 4 CMB and galaxy lensing surveys assuming a weak prior, respectively. Figures \ref{fig:paramconstraintstightcmbcmbprior} and \ref{fig:paramconstraintstightgalcmbprior} show confidence ellipses in the case of our primary CMB $T+E$ prior. Table \ref{tab:paramconstraintstight} summerizes all forecasted constraints.

@@

    \item For lower noise values (in particular for stage 4 noise) the bispectra become increasingly important compared to the powerspectra. The lensing bispectra are directly related to the amount of non-gaussianity in the matter distribution. As noise levels become lower, we are able to measure the matter distribution on small enough scales where nonlinear matter evolution becomes important. Additionally, for galaxy lensing, the window function peaks at later redshifts than compared to CMB lensing. The bispectrum is largest at late times, and thus the galaxy lensing bispectrum is significantly easier to detect and is better at constraining parameters then the CMB lensing bispectrum.

@@

    \item Stage 4 galaxy weak lensing surveys generally contain more information than CMB surveys. 
    % Comparing the confidence ellipses, it is clear that the two types of surveys depend on the underlying parameters in different ways. This leads to substantially better constraints for $\sigma_8$, $\Omega_m$, $m_\nu$, and $w_0$ when combining CMB and galaxy lensing.
\end{itemize}

\begin{table}
    \centering
    \scriptsize
    \begin{tabular}{|l|r|rrr|rrr|rrr|}
        \hline
        \multicolumn{11}{|c|}{weak priors} \\
        \hline
        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
        \hline
        Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
        \hline
        $m_\nu$    &     1.00e+01 &  2.82e-01 &  1.13e+00 &        2.26e-01 & 2.60e-01 & 4.07e-01 &      1.85e-01 &      9.54e-02 &      1.81e-01 &             6.46e-02 \\
        $w_0$      &     1.00e+01 &  5.42e-01 &  1.44e+00 &        3.87e-01 & 2.14e-01 & 1.28e-01 &      3.22e-02 &      1.51e-01 &      1.13e-01 &             2.45e-02 \\
        $\sigma_8$ &     2.46e+00 &  1.18e-01 &  1.95e-01 &        5.69e-02 & 5.36e-02 & 5.43e-02 &      1.73e-02 &      3.48e-02 &      3.32e-02 &             9.20e-03 \\
        $\Omega_m$ &     6.70e-01 &  1.95e-01 &  1.84e-01 &        7.73e-02 & 7.57e-02 & 5.88e-02 &      1.81e-02 &      5.15e-02 &      3.76e-02 &             8.57e-03 \\
       \hline
        \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
        \hline
        && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
        \hline
         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
        \hline
        $m_\nu$    &    2.16e-01 &  5.28e-02 &  1.21e-01 &        5.22e-02 & 1.48e-01 & 1.49e-01 &      1.12e-01 &      1.88e-02 &      1.02e-01 &             1.62e-02 \\
        $w_0$      &    5.78e-02 &  4.19e-02 &  5.30e-02 &        4.19e-02 & 4.66e-02 & 4.77e-02 &      2.55e-02 &      3.35e-02 &      4.33e-02 &             1.99e-02 \\
        $\sigma_8$ &    1.55e-02 &  1.06e-02 &  1.50e-02 &        1.06e-02 & 1.18e-02 & 1.21e-02 &      8.18e-03 &      8.99e-03 &      1.19e-02 &             5.28e-03 \\
        $\Omega_m$ &    1.50e-02 &  1.20e-02 &  1.46e-02 &        1.20e-02 & 1.34e-02 & 1.23e-02 &      9.32e-03 &      1.13e-02 &      1.19e-02 &             6.16e-03 \\
       \hline
        \end{tabular}
\caption{Parameter constraints for different combinations of weak lensing information. $C$ stands for powerspectra, and $B$ for bispectra. All surveys are assumed to be stage 4 surveys with multipole range $2 \leq \ell \leq 2000$. We use the weak and CMB priors described in table \ref{tab:fiducialpars}.}
\label{tab:paramconstraintstight}
\end{table}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb_weak_prior.pdf}
    \caption{Parameter constraints and confidence ellipses for $\mu_\nu$, $w_0$, $\sigma_8$, and $\Omega_m$. Using ``stage 4'' noise with multipole range from $2$ to $2000$. The colored plots show CMB lensing power- and/or bispectrum constraints. We use the weak priors listed in table \ref{tab:fiducialpars}. The confidence ellipses are for $1\sigma$ certainty. They show approximate degeneracies in the information gained from a survey if they are ``stretched''. When the information of ellipses with degeneracies in different directions is combined, the degeneracies are removed and the constraints typically becomes much better on the relevant parameters. The black plots show the constraints using all lensing information, i.e. CMB lensing spectra, galaxy lensing spectram, and all cross spectra.}
    \label{fig:paramconstraintstightcmb}
\end{figure}

@@

\section{Discussion and conclusion}\label{sec:discussion}
In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power- and bispectra, considering experimental parameters representative of Stage 3 and Stage 4 CMB and galaxy surveys. Our analysis shows that while lensing power spectra are detectable at high significance, the inclusion of bispectra generally offers significant improvements when considering parameter constraints when considering an LCDM+ model of the universe. Notably, both CMB and galaxy bispectra are found to be detectable, even with Stage 3 experimental noise levels.

The primary impact of incorporating bispectrum information is a tightening of cosmological parameter constraints, particularly for parameters such as $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies. We found that for sufficiently low noise levels, characteristic of Stage 4 surveys, the bispectra themselves can offer constraining power comparable to that of the power spectra, due to nonlinear structure formation on small scales being better detected by these future surveys. Furthermore, the combination of CMB and galaxy lensing probes, especially when both power spectra and bispectra are utilized, further enhances constraints and aids in mitigating degeneracies. However, lensing power spectra alone, even for Stage 4 surveys, appear insufficient to competitively constrain an 8-parameter $\Lambda\text{CDM} + w_0 + \sum m_\nu$ model without priors.

@@

     \item We assumed throughout this paper that surveys would measure multipoles as low as $\ell = 2$.  However, our analytical formulas to calculate lensing spectra utilize the Limber approximation, which is only robust for $\ell \gtrsim 50$. Avoiding the Limber approximation would complicate calculations, requiring knowledge of unequal-time matter power and bispectra, and would be computationally much more demanding due to the need to evaluate double integrals instead of single integrals. With the methods used in this paper, recalculating Fisher matrices without the Limber approximation would not be feasible, even if unequal-time matter spectra were readily available. Methods such as those explored in reference \cite{Chen:2021vba} might offer a path to address this.

>>>>

\usepackage[utf8]{inputenc}   % Allows UTF-8 character encoding
\usepackage{amsmath}          % For advanced math typesetting
\usepackage{amssymb}          % Additional math symbols
\usepackage{graphicx}         % For including images
\usepackage{hyperref}         % For hyperlinks in the document
\usepackage{geometry}         % For setting up page geometry
\geometry{margin=1in}         % 1-inch margins all around
\usepackage{siunitx}          % For consistent typesetting of units
\usepackage{float}            % To control float positioning (e.g., figures/tables)
\usepackage{enumitem}         % Better control over list formatting
\usepackage{tocloft}
\usepackage{tikz}
\usepackage{tikz-3dplot}
\usetikzlibrary{shapes.geometric, arrows, backgrounds}
\usepackage{empheq}
\usepackage{booktabs}
\usepackage{multirow} % needed for merging cells vertically
\usepackage{booktabs}
\usepackage[table]{xcolor}

@@

\emailAdd{jonasfrugte@gmail.com}
\emailAdd{p.d.meerburg@rug.nl}
\abstract{
Upcoming stage 4 surveys, such as the Simons Observatory, LSST, and Euclid are poised to measure weak gravitational lensing of the Cosmic Microwave Background (CMB) and galaxies with unprecedented precision. While the power spectrum is the standard statistic used to analyze weak lensing data, non-Gaussianity induced by the non-linear growth of structure imparts significant cosmological information into higher-order statistics. In this paper, we forecast the ability of future surveys to constrain cosmological parameters using the weak lensing power spectrum and bispectrum from both CMB and galaxy surveys, as well as their cross-correlations. We consider an eight-parameter cosmological model ($\Lambda$CDM + $\sum m_{\nu}$ + $w_{0}$) and assess parameter constraints for stage 4 survey specifications. Our analysis shows that, in the absence of systematics, both the CMB and galaxy lensing bispectra will be detectable at high signal-to-noise. We consider parameter constraints with a ``weak'' prior and with a stronger prior from the primary CMB anisotropies. In the weak prior case the lensing bispectrum substantially improves parameter constraints over using the lensing power spectrum alone, primarily by breaking parameter degeneracies. For the stronger prior we see less significant gains, particularly for the CMB lensing bispectrum. We further find that for stage 4 surveys that are able to measure lensing on small scales where non-linear structure formation becomes important, the constraining power of the lensing bispectrum can become comparable to that of the lensing power spectrum. Furthermore, we demonstrate that, both for our weak and strong priors, the synergy between CMB and galaxy lensing is substantial. Combining the two probes, particularly with both their lensing power- and bi-spectra, leads to significantly tighter constraints than achievable with either probe individually, with the greatest improvement seen for the sum of neutrino masses. Additionally, we also analyze the changes in detectability and parameter constraints when taking into account post-Born corrections to the lensing bispectra, as these were shown to be significant, especially for the CMB lensing bispectrum \cite{postborn_pratten_lewis}. We show that the conclusions of this paper remain largely the same with this correction.
}

@@

Weak lensing has been used as a probe to constrain cosmological parameters for over a decade \cite{kilbinger2015cosmology,Bacon2000,Kaiser2000,Waerbeke2000,Bartelmann2001,Planck2018Lensing}. A range of upcoming surveys, such as the survey by the Simons Observatory (SO) \cite{Ade2019}, the Legacy Survey of Space and Time (LSST) \cite{Ivezic2019}, and Euclid \cite{Laureijs2011} aims to measure weak lensing of the cosmic microwave background (CMB) and galaxies respectively. These upcoming surveys are expected to achieve significantly higher accuracy than their predecessors, leading to tighter constraints on cosmological parameters and advancing our overall understanding of cosmology. Typically, the standard summary statistic of interest is the power spectrum (or equivalently, two-point function) of the lensing potential (as well as cross-correlations with galaxy clustering). This can be measured in terms of the lensing convergence $\kappa$, lensing shear $\gamma$, or lensing potential $\psi$. In the weak lensing regime and under the Born approximation, they are all equivalent, in the sense that they can be directly converted into one another (appendix \ref{sec:weaklensstats}). In CMB surveys, the temperature and polarization anisotropies are used to reconstruct the lensing potential \cite{Hu_2002}, while with galaxies, one instead calculates lensing effects by measuring the lensing shear, which can be deduced from the ellipticity distribution of the observed galaxies \cite{HoekstraJain2008}.

Even with Gaussian initial conditions, there are significant amounts of non-Gaussianity in the weak lensing signal, especially in galaxy lensing due to it being a probe of the more recent distribution of matter \cite{Bernardeau1997,Takada2003}. In addition to looking at the power spectrum, a natural next step is to measure weak lensing bispectra. While the bispectrum has been measured for galaxy lensing signals \cite{vanWaerbeke2002}, a definitive detection of the CMB lensing bispectrum is still pending \cite{Kalaja_2023,Namikawa_2016}. Upcoming surveys, such as the ones mentioned earlier, may be able to detect this signal. Therefore it is timely to consider the potential of higher-order spectra to constrain cosmological parameters.

@@

\subsection{Parameter Constraints}
Next, we consider cosmological parameter constraints. We include both the power spectrum and the bispectrum of the weak lensing. We look at two extension of $\Lambda CDM$, $m_\nu$ and $w_0$ as well as the derived parameters $\sigma_8$, and $\Omega_m$. Constraints on the full set of $\Lambda$CDM parameters can be found in appendix \ref{sec:lambdacdmconstraints}. Results shown include marginalization over all parameters. Figures \ref{fig:paramconstraintstightcmb} and \ref{fig:paramconstraintstightgal} show confidence ellipses assuming a weak prior from stage 4 CMB and galaxy lensing surveys, respectively. Figures \ref{fig:paramconstraintstightcmbcmbprior} and \ref{fig:paramconstraintstightgalcmbprior} similarly show confidence ellipses in the case of our primary CMB $T+E$ prior. Table \ref{tab:paramconstraintstight} summarizes all forecasted constraints.

@@

    \item For lower noise values (in particular for stage 4 noise), the bispectra become increasingly important compared to the power spectra. The lensing bispectra are directly related to the amount of non-Gaussianity in the matter distribution. As noise levels become lower, we are able to measure the matter distribution on small enough scales where nonlinear matter evolution becomes important. Additionally, for galaxy lensing, the window function peaks at later redshifts compared to CMB lensing. The bispectrum is largest at late times, and thus the galaxy lensing bispectrum is significantly easier to detect and is better at constraining parameters than the CMB lensing bispectrum.

@@

    \item Stage 4 galaxy weak lensing surveys generally contain more information than CMB surveys. 
    Comparing the confidence ellipses, it is clear that the two types of surveys depend on the underlying parameters in different ways. This leads to substantially better constraints for $\sigma_8$, $\Omega_m$, $m_\nu$, and $w_0$ when combining CMB and galaxy lensing.

    \item It is clear that post-Born corrections do not change the constraints drastically. As expected, they matter less for galaxy lensing and when using the CMB $T+E$ prior. Interestingly, although the constraints due to bispectra alone are always worse, for some parameters, they are instead improved when combining lensing bispectrum information with that of the lensing power spectrum. This implies that, at least in some cases, the post-Born contribution to the lensing bispectrum depends on the parameters in a different way than the lensing power spectrum and thus improves the amount of information we can extract from survey data despite making the lensing bispectrum less detectable. Besides these observations, we conclude that taking into account post-Born corrections does not lead to different conclusions about the overall sensitivity of the lensing spectra.
\end{itemize}

\begin{table}
    \centering
    \scriptsize
%     \begin{tabular}{|l|r|rrr|rrr|rrr|}
%         \hline
%         \multicolumn{11}{|c|}{weak priors} \\
%         \hline
%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
%         \hline
%         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
% \hline
%  $m_\nu$    & 1.000e+01 &  2.815e-01 &                  1.104e+00 &                           2.247e-01 &  9.954e-01 &                  2.646e-01 &                           1.551e-01 &  1.042e-01 &                  1.317e-01 &                           6.537e-02 \\
%  $w_0$      & 1.000e+01 &  5.415e-01 &                  1.405e+00 &                           3.683e-01 &  2.638e-01 &                  7.817e-02 &                           2.475e-02 &  5.936e-02 &                  6.629e-02 &                           1.899e-02 \\
%  $\sigma_8$ & 2.457e+00 &  1.181e-01 &                  1.872e-01 &                           5.486e-02 &  8.069e-02 &                  3.668e-02 &                           1.299e-02 &  1.375e-02 &                  1.587e-02 &                           8.048e-03 \\
%  $\Omega_m$ & 6.702e-01 &  1.948e-01 &                  1.790e-01 &                           7.270e-02 &  1.434e-01 &                  4.831e-02 &                           1.667e-02 &  1.619e-02 &                  1.607e-02 &                           7.233e-03 \\
%        \hline
%         \multicolumn{11}{|c|}{CMB $T + E$ priors} \\
%         \hline
%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
%         \hline
%          Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
%         \hline
%  $m_\nu$    & 2.157e-01 &  7.445e-02 &                  1.178e-01 &                           7.283e-02 &  1.406e-01 &                  1.042e-01 &                           8.592e-02 &  4.340e-02 &                  8.965e-02 &                           3.735e-02 \\
%  $w_0$      & 5.781e-02 &  4.606e-02 &                  5.285e-02 &                           4.578e-02 &  4.756e-02 &                  3.934e-02 &                           2.080e-02 &  3.231e-02 &                  3.667e-02 &                           1.741e-02 \\
%  $\sigma_8$ & 1.547e-02 &  1.360e-02 &                  1.492e-02 &                           1.349e-02 &  1.252e-02 &                  1.136e-02 &                           7.881e-03 &  9.682e-03 &                  1.023e-02 &                           6.126e-03 \\
%  $\Omega_m$ & 1.495e-02 &  1.381e-02 &                  1.457e-02 &                           1.376e-02 &  1.336e-02 &                  1.124e-02 &                           8.715e-03 &  1.028e-02 &                  9.616e-03 &                           5.484e-03 \\
%        \hline
%             \multicolumn{11}{|c|}{weak priors, with post-Born corrections} \\
%         \hline
%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
%         \hline
%         Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
% \hline
% $m_\nu$    & 1.000e+01 &  2.815e-01 &                  1.158e+00 &                           1.805e-01 &  9.954e-01 &                  3.349e-01 &                           2.073e-01 &  1.042e-01 &                  1.775e-01 &                           6.256e-02 \\
%  $w_0$      & 1.000e+01 &  5.415e-01 &                  8.193e-01 &                           3.051e-01 &  2.638e-01 &                  8.328e-02 &                           2.825e-02 &  5.936e-02 &                  6.271e-02 &                           1.988e-02 \\
%  $\sigma_8$ & 2.457e+00 &  1.181e-01 &                  1.846e-01 &                           5.660e-02 &  8.069e-02 &                  4.047e-02 &                           1.317e-02 &  1.375e-02 &                  1.782e-02 &                           8.369e-03 \\
%  $\Omega_m$ & 6.702e-01 &  1.948e-01 &                  1.810e-01 &                           7.086e-02 &  1.434e-01 &                  5.474e-02 &                           1.811e-02 &  1.619e-02 &                  1.799e-02 &                           7.325e-03 \\
%        \hline
%         \multicolumn{11}{|c|}{CMB $T + E$ priors, with post-Born corrections} \\
%         \hline
%         && \multicolumn{3}{c|}{CMB lensing}& \multicolumn{3}{c|}{Gal. lensing}& \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
%         \hline
%          Par            &   prior &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ &   $C_\ell$ &   $B_{\ell_1\ell_2\ell_3}$ &   $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
%         \hline
%  $m_\nu$    & 2.157e-01 &  7.445e-02 &                  1.298e-01 &                           7.313e-02 &  1.406e-01 &                  1.166e-01 &                           9.869e-02 &  4.340e-02 &                  1.034e-01 &                           3.766e-02 \\
%  $w_0$      & 5.781e-02 &  4.606e-02 &                  5.308e-02 &                           4.596e-02 &  4.756e-02 &                  4.163e-02 &                           2.220e-02 &  3.231e-02 &                  3.911e-02 &                           1.798e-02 \\
%  $\sigma_8$ & 1.547e-02 &  1.360e-02 &                  1.498e-02 &                           1.354e-02 &  1.252e-02 &                  1.168e-02 &                           7.717e-03 &  9.682e-03 &                  1.073e-02 &                           6.285e-03 \\
%  $\Omega_m$ & 1.495e-02 &  1.381e-02 &                  1.467e-02 &                           1.379e-02 &  1.336e-02 &                  1.141e-02 &                           8.530e-03 &  1.028e-02 &                  1.017e-02 &                           5.687e-03 \\
%        \hline
%         \end{tabular}
 \begin{tabular}{|l|r|rrr|rrr|rrr|}
        \hline
        \multicolumn{11}{|c|}{weak priors (with and without post-Born corrections)} \\
        \hline
        && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
        \hline
        Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
        \hline
$m_\nu$    & 10 & 0.28 & 1.1 & 0.23 & 1.0 & 0.26 & 0.16 & 0.10 & 0.13 & 0.065 \\
\rowcolor{blue!10}
           &    &  & 1.2 & 0.18 & & 0.33 & 0.21 & & 0.18 & 0.063 \\
$w_0$      & 10 & 0.54 & 1.4 & 0.37 & 0.26 & 0.078 & 0.025 & 0.059 & 0.066 & 0.019 \\
\rowcolor{blue!10}
           &    &  & 0.82 & 0.31 &  & 0.083 & 0.028 & & 0.063 & 0.020 \\
$\sigma_8$ & 2.5 & 0.12 & 0.19 & 0.055 & 0.081 & 0.037 & 0.013 & 0.014 & 0.016 & 0.0080 \\
\rowcolor{blue!10}
           &     &  & 0.18 & 0.057 & & 0.040 & 0.013 & & 0.018 & 0.0084 \\
$\Omega_m$ & 0.67 & 0.19 & 0.18 & 0.073 & 0.14 & 0.048 & 0.017 & 0.016 & 0.016 & 0.0072 \\
\rowcolor{blue!10}
           &     &  & 0.18 & 0.071 & & 0.055 & 0.018 & & 0.018 & 0.0073 \\
        \hline
        \multicolumn{11}{|c|}{CMB $T + E$ priors (with and without post-Born corrections)} \\
        \hline
        && \multicolumn{3}{c|}{CMB lensing} & \multicolumn{3}{c|}{Gal. lensing} & \multicolumn{3}{c|}{CMB $\times$ Gal. lensing} \\
        \hline
        Par & prior & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ & $C_\ell$ & $B_{\ell_1\ell_2\ell_3}$ & $C_\ell + B_{\ell_1\ell_2\ell_3}$ \\
        \hline
$m_\nu$    & 0.22 & 0.074 & 0.12 & 0.073 & 0.14 & 0.10 & 0.086 & 0.043 & 0.090 & 0.037 \\
\rowcolor{blue!10}
           &      & & 0.13 & 0.073 & & 0.12 & 0.099 & & 0.10 & 0.038 \\
$w_0$      & 0.058 & 0.046 & 0.053 & 0.046 & 0.048 & 0.039 & 0.021 & 0.032 & 0.037 & 0.017 \\
\rowcolor{blue!10}
           &       & & 0.053 & 0.046 & & 0.042 & 0.022 &  & 0.039 & 0.018 \\
$\sigma_8$ & 0.015 & 0.014 & 0.015 & 0.013 & 0.013 & 0.011 & 0.0079 & 0.0097 & 0.010 & 0.0061 \\
\rowcolor{blue!10}
           &       & & 0.015 & 0.014 &  & 0.012 & 0.0077 & & 0.011 & 0.0063 \\
$\Omega_m$ & 0.015 & 0.014 & 0.015 & 0.014 & 0.013 & 0.011 & 0.0087 & 0.010 & 0.0096 & 0.0055 \\
\rowcolor{blue!10}
           &       & & 0.015 & 0.014 & & 0.011 & 0.0085 & & 0.010 & 0.0057 \\
        \hline
    \end{tabular}

\caption{Parameter constraints for different combinations of weak lensing information. $C$ stands for power spectra, and $B$ for bispectra. All surveys are assumed to be stage 4 surveys with multipole range $2 \leq \ell \leq 2000$. We use the weak and CMB priors described in table \ref{tab:fiducialpars}. The blue colored rows show constraints from bispectra with post-Born corrections. Post-Born corrections are not taken into account for the lensing power spectra as mentioned earlier, so these values are omitted in the blue rows.}
\label{tab:paramconstraintstight}
\end{table}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/param_constraints_tight_cmb_weak_prior.pdf}
    \caption{Parameter constraints and confidence ellipses for $\mu_\nu$, $w_0$, $\sigma_8$, and $\Omega_m$. Using ``stage 4'' noise with multipole range from $2$ to $2000$. The colored plots show CMB lensing power and/or bispectrum constraints. We use the weak priors listed in table \ref{tab:fiducialpars}. The confidence ellipses are for $1\sigma$ certainty. They show approximate degeneracies in the information gained from a survey if they are ``stretched''. When the information of ellipses with degeneracies in different directions is combined, the degeneracies are removed and the constraints typically become much better on the relevant parameters. The black plots show the constraints using all lensing information, i.e. CMB lensing spectra, galaxy lensing spectra, and all cross spectra.}
    \label{fig:paramconstraintstightcmb}
\end{figure}

@@

\section{Discussion and conclusion}\label{sec:discussion}
In this work, we have presented parameter forecasts from CMB lensing and galaxy lensing power and bispectra, considering experimental parameters representative of stage 3 and stage 4 CMB and galaxy surveys. Our analysis shows that while lensing power spectra are detectable at high significance, the inclusion of bispectra generally offers significant improvements when considering parameter constraints when considering an LCDM+ model of the universe. Notably, both CMB and galaxy bispectra are found to be detectable, even with stage 3 experimental noise levels.

The primary impact of incorporating bispectrum information is a tightening of cosmological parameter constraints, particularly for parameters such as $H_0$, $\sigma_8$, and $\Omega_m$, and in breaking key parameter degeneracies. We found that for sufficiently low noise levels, characteristic of stage 4 surveys, the bispectra themselves can offer constraining power comparable to that of the power spectra, due to nonlinear structure formation on small scales being better detected by these future surveys. Furthermore, the combination of CMB and galaxy lensing probes, especially when both power spectra and bispectra are utilized, further enhances constraints and aids in mitigating degeneracies. If all information were to be combined\footnote{And using the CMB $T+E$ prior.} we find particularly tight constraints on neutrino mass and dark energy equation of state in particular, i.e. $\sigma(m_\nu) = 55$ meV and $\sigma(w_0) = 0.017$. However, lensing power spectra alone, even for stage 4 surveys, appear insufficient to competitively constrain an 8-parameter $\Lambda\text{CDM} + w_0 + \sum m_\nu$ model without any priors.

@@

     \item We assumed throughout this paper that lensing surveys would use data from multipoles as low as $\ell = 2$.  However, our analytical formulas to calculate lensing spectra utilize the Limber approximation, which is only robust for $\ell \gtrsim 50$. Avoiding the Limber approximation would complicate calculations, requiring knowledge of unequal-time matter power and bispectra\footnote{Technically, this would be computationally much more demanding due to the need to evaluate double integrals instead of single integrals}. Moving beyond the Limber approximation can be achieved using methods such as those explored in reference \cite{Chen:2021vba} and might offer a path to address this.
